---
title: "spongeDNA_ob1_code"
author: "cai wang"
date: "20/12/2021"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(phyloseq)
library(ggplot2)
library(data.table)
library(sjmisc)
library(vegan)
library(pheatmap)
library(RColorBrewer)
library(car)
library(ggrepel)
library(mvabund)
library(HH)
library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

#######filter: filtering OTU with reads <10
according to positive control, The smallest read is 11, so we select 10 as the threshold to filter "OTU[OTU <= 10] <- 0"and we plot all control samples with unfiltered OTU and reads.#####

before filter we do two more formative steps:
one: transfer data to phyloseq format
```{r phyloseq formal}
OTUseq<-read.csv("spongetank_SWARM3_OTU_LULU_20220220_collapsed.csv", header=TRUE)
#otu_table
names(OTUseq)
OTU<-OTUseq %>% select (D1.SW.BLANK:LIB02.PCR.Pos.03)
rownames(OTU)=OTUseq$id

#tax_table
TAX<-OTUseq %>% select (kingdom_name:species_name)
rownames(TAX)=OTUseq$id
TAX <- as.matrix(TAX)

#sample_data
sample_data<-read.csv("sampledata_full_20211207.csv", header=TRUE)
sample_data=arrange(sample_data,name)
otusum=data.frame(name=names(OTU),quant_reading=colSums(OTU))
otusum=arrange(otusum,name)
sample_data=left_join(sample_data,otusum,by="name")
rownames(sample_data)=sample_data$name

#####filter total reads=0 samples
sample_data=sample_data %>% filter(quant_reading!=0)
OTU= OTU %>% select(sample_data$name)
OTU <- as.matrix(OTU)
###write the phyloseq format
physeq_spongetank = phyloseq(otu_table(OTU, taxa_are_rows = TRUE), tax_table(TAX), sample_data(sample_data))

test=data.frame(otu_table(physeq_spongetank))
colSums(test)
save(physeq_spongetank,file = "physeq_spongetank_20220220.rdata")
```
new remove sequence in each sample which reads abundance <10
```{r}
load("physeq_spongetank_20220220.rdata")
OTU<- data.frame(t(as.data.frame(otu_table(physeq_spongetank))))
metadata<- data.frame(sample_data(physeq_spongetank)) 
TAX=data.frame(tax_table(physeq_spongetank))
s.reads<-sort(apply(OTU,1,sum))#13331.82, 6192.966
sd(s.reads);mean(s.reads)
o.reads<-sort(apply(OTU,2,sum))#29814.04, 10021.35
sd(o.reads); mean(o.reads)
OTUname=data.frame(OTU=rownames(TAX),family=TAX$family_name,genus=TAX$genus_name,species=TAX$species_name)
OTUname$species=as.character(OTUname$species)
OTUname$OTU=as.character(OTUname$OTU)
OTUname=raster::trim(OTUname)
OTU[OTU <= 10] <- 0
exclude.OTU = which(colSums(OTU) == 0)
OTU= OTU[,-exclude.OTU]
OTUselect= data.frame(OTU=colnames(OTU))
OTUname=left_join(OTUselect,OTUname,by="OTU")
#we add species name to OTU table.
#tax_name correction
for (i in 1:length(OTUname$species)) {
  if(OTUname$species[i]=="")
    OTUname$species[i] = paste0(OTUname$genus[i]," ","sp" )
  if(OTUname$genus[i]=="")
    OTUname$species[i] = paste0("f_",OTUname$family[i],"sp" )
}
for (i in 1:length(OTUname$species)) {
  if(OTUname$species[i]=="f_sp")
    OTUname$species[i] = paste0("OTU",i)
}
OTUname$species=gsub(" ", "_", OTUname$species)

TAX$OTU=rownames(TAX)
TAX=left_join(OTUname,TAX,by="OTU")
#collapse two Chromis and Zebrasoma sp into one otu
names(OTU)
OTU <- OTU %>% mutate(Chromis_viridis = 	tank_000000102 + tank_000000006) 
OTU <- OTU %>% select(-tank_000000102,-tank_000000006)
OTU <- OTU %>% mutate(Zebrasoma_veliferum = tank_000000009 + tank_000000820) 
OTU <- OTU %>% select(-tank_000000009,-tank_000000820)
OTU <- rotate_df(OTU)
OTU$OTU=rownames(OTU)
OTU2=left_join(OTU,OTUname,by="OTU")
OTU2$species[60]=OTU2$OTU[60]
OTU2$species[61]=OTU2$OTU[61]
OTU2=OTU_before_clean
rownames(OTU2)=OTU2$species
rownamesOTU2=rownames(OTU2)
OTU2 <- OTU2 %>% select(-OTU,-family,-species,-genus)
OTU2<- apply(OTU2,2,as.numeric)
rownames(OTU2)=rownamesOTU2
```
pheatmap, before filtering small reads, separate and plot the controls and samples
```{r}
OTU_before_clean <- read_csv("OTU_before_clean.csv")
OTU3=data.frame(t(OTU_before_clean))
OTU.sam=OTU3 %>% filter(metadata$hour!="env" & metadata$hour!="CONTROL" & metadata$hour!="Pos")
OTU.ctl=OTU3 %>% filter(metadata$hour=="env" | metadata$hour=="CONTROL" | metadata$hour=="Pos")
metadata.ctl=metadata %>% filter(hour=="env" | hour=="CONTROL" | hour=="Pos")
metadata.sam=metadata %>% filter(hour!="env" & hour!="CONTROL" & hour!="Pos")
communityphm<-OTU.ctl
communityphm$control<-metadata.ctl$tank
communityphm<-arrange(communityphm, control)
metadata.ctl<-arrange(metadata.ctl, tank)
communityphm <- communityphm %>% dplyr::select(-control)
communityphm2=communityphm
#remove small reads and plot again
communityphm2=log(communityphm2)
communityphm2[communityphm2 == "-Inf"] <- 0
exclude2 = which(colSums(communityphm2) == 0)
communityphm2=communityphm2[,-exclude2]
#set group
colvec <- brewer.pal(9, "Blues")
colvec <-c("white",colvec,"#053061") 
ann_colors<- brewer.pal(8, "Set1")
pheatmap(communityphm2,cluster_rows = F,cluster_cols = F,border_color = NA,color= colvec, clustering_method="ward.D2",fontsize_row=8,margins=c(5,10))

rm(communityphm,communityphm2)
####################set group for sample
communityphm<-OTU.sam
communityphm$control<-metadata.sam$tank
communityphm<-arrange(communityphm, control)
metadata.sam<-arrange(metadata.sam, tank)
communityphm <- communityphm %>% dplyr::select(-control)
communityphm2=communityphm
communityphm=log(communityphm)
communityphm[communityphm == "-Inf"] <- 0
exclude = which(colSums(communityphm) == 0)
communityphm=communityphm[,-exclude]
colvec <- brewer.pal(9, "Blues")
colvec <-c("white",colvec,"#053061") 
pheatmap(communityphm,cluster_rows = F,cluster_cols = F,border_color = NA,color= colvec, clustering_method="ward.D2",fontsize_row=8,margins=c(5,10))
rm(communityphm,communityphm2)
```

```{r}
##TAX$OTU=rownames(TAX)
TAX <- TAX %>% filter(OTU!="tank_000000006" & OTU!="tank_000000009")
TAX$species[14]="Zebrasoma_veliferum"
rownames(TAX)=TAX$species
TAX <- TAX %>% select(-OTU)
TAX$genus_name=as.character(TAX$genus_name)
TAX$family_name=as.character(TAX$family_name)

TAX=raster::trim(TAX)
for (i in 1:length(TAX$genus_name)) {
  if(TAX$genus_name[i]=="")
    TAX$genus_name[i] = paste0("unassigned")
  if(TAX$family_name[i]=="")
    TAX$family_name[i] = paste0("unassigned")
}
for (i in 1:length(TAX$species_name)) {
  if(TAX$genus_name[i]=="unassigned")
    TAX$species_name[i]=paste0("unassigned")
}

TAX=as.matrix(TAX)
physeq_spongetank = phyloseq(otu_table(OTU2, taxa_are_rows = TRUE), tax_table(TAX), sample_data(metadata))
save(physeq_spongetank,file = "physeq_spongetank_clean_20220220.rdata")
```
after filtering small reads, separate and plot the controls and samples
```{r pheatmap}
rm(list=ls())
load("physeq_spongetank_clean_20220220.rdata")
OTU<- data.frame(t(as.data.frame(otu_table(physeq_spongetank))))
#summary the total number of reads
summary1=data.frame(name=names(OTU),sum=colSums(OTU))
write_csv(summary1, "OTU_read_summary.csv")
metadata<- data.frame(sample_data(physeq_spongetank))
OTU.sam=OTU %>% filter(metadata$hour!="env" & metadata$hour!="CONTROL" & metadata$hour!="Pos")
OTU.ctl=OTU %>% filter(metadata$hour=="env" | metadata$hour=="CONTROL" | metadata$hour=="Pos")
metadata.ctl=metadata %>% filter(hour=="env" | hour=="CONTROL" | hour=="Pos")
metadata.sam=metadata %>% filter(hour!="env" & hour!="CONTROL" & hour!="Pos")


communityphm<-OTU.ctl
communityphm$control<-metadata.ctl$tank
communityphm<-arrange(communityphm, control)
metadata.ctl<-arrange(metadata.ctl, tank)
communityphm <- communityphm %>% dplyr::select(-control)
communityphm2=communityphm
#remove small reads and plot again
communityphm2=log(communityphm2)
communityphm2[communityphm2 == "-Inf"] <- 0
exclude2 = which(colSums(communityphm2) == 0)
communityphm2=communityphm2[,-exclude2]
or<-order(names(communityphm2))
names(communityphm2)
or<-c(1:5,7,24,10,6,8,9,11:23)
communityphm2<-communityphm2[,or]
#set group
colvec <- brewer.pal(9, "Blues")
colvec <-c("white",colvec,"#053061") 
ann_colors<- brewer.pal(8, "Set1")
# pheatmap(communityphm2,annotation_colors = ann_colors,annotation_row = annotation_row,cluster_rows = F,cluster_cols = F,border_color = NA,color= colvec, clustering_method="ward.D2",fontsize_row=8,margins=c(5,10))
pheatmap(communityphm2,cluster_rows = F,cluster_cols = F,border_color = NA,color= colvec, clustering_method="ward.D2",fontsize_row=8,margins=c(5,10))

rm(communityphm,communityphm2)
####################set group for sample
communityphm<-OTU.sam
communityphm$control<-metadata.sam$tank
communityphm<-arrange(communityphm, control)
metadata.sam<-arrange(metadata.sam, tank)
communityphm <- communityphm %>% dplyr::select(-control)
communityphm2=communityphm
communityphm=log(communityphm)
communityphm[communityphm == "-Inf"] <- 0
exclude = which(colSums(communityphm) == 0)
communityphm=communityphm[,-exclude]
# or<-order(names(communityphm))
# names(communityphm)
# or<-c(1,2,4,6,7,9,10:12,16,24,25,26,51,52,3,5,8,13:15,17:23,27,50)
# communityphm<-communityphm[,or]

colvec <- brewer.pal(9, "Blues")
colvec <-c("white",colvec,"#053061") 
ann_colors<- brewer.pal(8, "Set1")
pheatmap(communityphm,cluster_rows = F,cluster_cols = F,border_color = NA,color= colvec, clustering_method="ward.D2",fontsize_row=8,margins=c(5,10))
```

#######filter: here we only keep targeted species for following analysis#####
filtering OTU that show in "day1 sponge samples", which are carried into the tanks by sponges, one OTU (contamination) show in sterivex filter blank, and all unassigned OTUs.
```{r third filter}
rm(list=ls())
load("physeq_spongetank_clean_20220220.rdata")
OTU<- data.frame(t(as.data.frame(otu_table(physeq_spongetank))))
metadata<- data.frame(sample_data(physeq_spongetank))
OTU.sam=OTU %>% filter(metadata$hour!="env" & metadata$hour!="CONTROL" & metadata$hour!="Pos")
exclude = which(colSums(OTU.sam) == 0)
OTU.sam=OTU.sam[,-exclude]
#remove positive control and human
names(OTU.sam)
colnames(OTU.sam)[2]="Hippocampus_semispinosis"
colnames(OTU.sam)[4]="Amphiprion_ocellaris"

OTU.sam=OTU.sam %>% select(Amphiprion_ocellaris,Chromis_viridis,Zebrasoma_veliferum,Hippocampus_semispinosis,Gramma_loreto)
TAX=data.frame(tax_table(physeq_spongetank))
OTU.sam$name=rownames(OTU.sam)
#i remove the contamination by hand, cause apparently the sponge brought it in from outside.
r1=which(OTU.sam$name == "D2.TANKB.FISH.SPONGE.PINK")
r2=which(OTU.sam$name == "D2.TANKC.FISH.SPONGE.CREAM")
OTU.sam$Zebrasoma_veliferum[r1]="0"
OTU.sam$Zebrasoma_veliferum[r2]="0"
OTU.sam$Zebrasoma_veliferum=as.numeric(OTU.sam$Zebrasoma_veliferum)
OTU.sam=OTU.sam %>% select(-name)
metadata.sam=metadata %>% filter(hour!="env" & hour!="CONTROL" & hour!="Pos")
OTU.sam=OTU.sam %>% filter(metadata.sam$sponge_type !="SP.BLACK")
metadata.sam=metadata.sam %>% filter(sponge_type !="SP.BLACK")
save(OTU.sam,metadata.sam,TAX,file = "sample_data_only_20220220.rdata")
```

#######statistical analysis start here##########
Q1:Does reads relate to detection rate or species size?
```{r load and format data}
rm(list=ls())
load("sample_data_only_20220220.rdata")
OTU.sam=log(OTU.sam)
OTU.sam[OTU.sam == "-Inf"] <- 0
OTU.full.i=bind_cols(metadata.sam,OTU.sam)
#OTU.full.i=OTU.full.i %>% filter(sponge_type =="WATER" | sponge_type =="SP.PINK" | sponge_type =="SP.CREAM")
sample_data<-read.csv("sampledata_full_20211207.csv", header=TRUE)
#sample_data=arrange(sample_data,name)
sample_data=left_join(sample_data,OTU.full.i,by="name")
rownames(sample_data)=sample_data$name
sample_data=sample_data%>%filter(sponge_type.x=="SP.PINK" | sponge_type.x=="SP.BLACK" |sponge_type.x=="SP.CREAM" | sponge_type.x=="WATER")
sample_data[is.na(sample_data)]<-0
OTU.full.i=sample_data%>%select(tank.x:sponge_type.x,Amphiprion_ocellaris:Gramma_loreto)
names(OTU.full.i)[1:6]=c("tank","hour","hours_fish","date","time","sponge_type")

OTU.full.i <- OTU.full.i %>%
  mutate(
    hours_fish = case_when(
      hours_fish == "0" ~ "20",
      hours_fish == "20" ~ "40",
      hours_fish == "24" ~ "44",
      hours_fish == "28" ~ "48",
      hours_fish == "44" ~ "64",
      hours_fish == "68" ~ "88",
      hours_fish == "92" ~ "112",
      TRUE ~ as.character(hours_fish)
    )
  )
OTU.full.i <- OTU.full.i %>%
  mutate(
    hour = case_when(
      hour == "0" ~ "20",
      hour == "20" ~ "40",
      hour == "24" ~ "44",
      hour == "28" ~ "48",
      hour == "44" ~ "64",
      hour == "68" ~ "88",
      hour == "92" ~ "112",
      TRUE ~ as.character(hour)
    )
  )
OTU.full.i <- OTU.full.i %>%
  mutate(
    sponge_type = case_when(
      sponge_type == "SP.BLACK" ~ "BLACK",
      sponge_type == "SP.PINK" ~ "PINK",
      sponge_type == "SP.CREAM" ~ "CREAM",
      TRUE ~ as.character(sponge_type)
    )
  )
OTU.full.i$hours_fish=as.factor(OTU.full.i$hours_fish)
OTU.full.i$hour=as.numeric(OTU.full.i$hour)
OTU.full.i$tank=as.factor(as.character(OTU.full.i$tank))
OTU.full.i$hours_fish=factor(OTU.full.i$hours_fish,levels = c('20','40','44','48','64','88','112')) 
OTU.full.i$sponge_type=as.factor(as.character(OTU.full.i$sponge_type))

OTU.sam=OTU.full.i%>%select(Amphiprion_ocellaris:Gramma_loreto)
OTU.sam2=OTU.full.i%>%select(Amphiprion_ocellaris:Gramma_loreto)
OTU.sam2[OTU.sam2 > 0] <- 1
OTU.sam2<- data.frame(apply(OTU.sam2,2,as.numeric))
```

```{r}
speciesinfo.t=data.frame(detection=colSums(OTU.sam2),logreads=colSums(OTU.sam),
name=c("Amphiprion ocellaris","Chromis viridis","Zebrasoma veliferum","Hippocampus semispinosis","Gramma loreto"))

ggplot(data = speciesinfo.t,aes(x = detection, y = logreads,label=name))+geom_point()+theme_classic()+ylab("log(reads)")+geom_smooth(method = "lm",se=T)+geom_text_repel(size=2)+scale_x_continuous(limits=c(0,35),breaks=seq(5,35,5))
#seprate sample type,c("SP.CREAM","SP.PINK","WATER")
OTU.sam.p=OTU.sam %>% filter(OTU.full.i$sponge_type=="PINK")
OTU.sam.p2=OTU.sam.p
OTU.sam.p2[OTU.sam.p2 > 0] <- 1
speciesinfo.p=data.frame(detection=colSums(OTU.sam.p2),logreads=colSums(OTU.sam.p),
name=c("Amphiprion ocellaris","Chromis viridis","Zebrasoma veliferum","Hippocampus semispinosis","Gramma loreto"))
speciesinfo.p$sample="PINK"
OTU.sam.c=OTU.sam %>% filter(OTU.full.i$sponge_type=="CREAM")
OTU.sam.c2=OTU.sam.c
OTU.sam.c2[OTU.sam.c2 > 0] <- 1
speciesinfo.c=data.frame(detection=colSums(OTU.sam.c2),logreads=colSums(OTU.sam.c),
name=c("Amphiprion ocellaris","Chromis viridis","Zebrasoma veliferum","Hippocampus semispinosis","Gramma loreto"))
speciesinfo.c$sample="CREAM"
OTU.sam.w=OTU.sam %>% filter(OTU.full.i$sponge_type=="WATER")
OTU.sam.w2=OTU.sam.w
OTU.sam.w2[OTU.sam.w2 > 0] <- 1
speciesinfo.w=data.frame(detection=colSums(OTU.sam.w2),logreads=colSums(OTU.sam.w),
name=c("Amphiprion ocellaris","Chromis viridis","Zebrasoma veliferum","Hippocampus semispinosis","Gramma loreto"))
speciesinfo.w$sample="WATER"
speciesinfo=bind_rows(speciesinfo.w,speciesinfo.p,speciesinfo.c)
col=c("#FFB90F", "#8B008B", "#7AC5CD")
names(speciesinfo)=c("Detection","logreads","Species","Sample type")
#A: for all, Adjusted R-squared:  0.9727 p= 0.00125 ** high relation
#Q2:Does reads and detection rate relate to time, differ from type?

#total
OTU.full.i$sampleread=rowSums(OTU.sam)
p1=ggplot(data = OTU.full.i,aes(x = hour, y = sampleread,color=sponge_type))+geom_point()+theme_classic()+geom_smooth(method = "glm",se=F)+scale_color_manual(values=c("#8B7D6B", "#FFB90F", "#8B008B","#7AC5CD"))+ylab("Log(reads)")+xlab("Time (hour)")+scale_x_continuous( limits=c(20,112),breaks=c(20,40,44,48,64,88,112))+scale_y_continuous( limits=c(0,30))+ theme(legend.position = 'none')
p1
p2=ggplot(data = speciesinfo,aes(x = Detection, y = logreads,color=`Sample type`,shape=Species))+geom_point()+theme_classic()+ylab("Log(reads)")+xlab("Detection")+scale_x_continuous(limits=c(0,10),breaks=seq(0,10,1))+scale_color_manual(values=col)+guides(fill=guide_colorbar(lable=TRUE,lable.theme=element_text(size=30,face="italic",angle = 45)))+geom_text_repel(aes(Detection, logreads, label=Species,size=3),show.legend = FALSE)
p2
ggpubr::ggarrange(p1,p2,ncol=2,nrow=1,labels=c("A","B"),common.legend = F)

OTU.full.i2=OTU.full.i%>%filter(sponge_type!="BLACK")
summary(lm(OTU.full.i2$hour~OTU.full.i2$sampleread))
#rm(OTU.full.i2,speciesinfo,speciesinfo.c,speciesinfo.p,speciesinfo.w,speciesinfo.t)

#pink
metadata.sam.p=OTU.full.i %>% filter(sponge_type=="PINK")
summary(lm(metadata.sam.p$hour~metadata.sam.p$sampleread))
#cream
metadata.sam.c=OTU.full.i %>% filter(sponge_type=="CREAM")
summary(lm(metadata.sam.c$hour~metadata.sam.c$sampleread))
#water
metadata.sam.w=OTU.full.i %>% filter(sponge_type=="WATER")
summary(lm(metadata.sam.w$hour~metadata.sam.w$sampleread))
TukeyHSD(aov(OTU.full.i$sampleread~OTU.full.i$sponge_type))
```
A:cream and water total reads decrease along time, but pink not.

Q3:what is the pattern of each sponge and water according to species richness and community
now look at α diversity
```{r incidence lm model}
OTU.full.i$incidence=rowSums(OTU.sam2)
p1=ggplot(OTU.full.i, aes(x = hour, y = incidence, color=sponge_type)) +geom_jitter(aes(colour=sponge_type),height=0.0,width =0.8,size=1.5)+geom_smooth(method = "lm",se=F)+ylab("Observed species richness")+xlab("Time (hour)")+scale_color_manual(values=c("#8B7D6B", "#FFB90F", "#8B008B","#7AC5CD"))+theme_bw()+theme(panel.grid=element_blank())+theme(legend.title=element_blank())+scale_x_continuous( limits=c(20,112),breaks=c(20,40,44,48,64,88,112))+scale_y_continuous( limits=c(0,3),breaks=c(0,1,2,3))
p2=ggplot(OTU.full.i, aes(x=factor(hours_fish), y=incidence,fill=sponge_type)) + 
  geom_boxplot() +
  geom_jitter(width =0.1,size=1,height=0)+ 
  theme_bw()+ylab("Observed species richness")+xlab("Time (hour)")+scale_fill_manual(values=c("#8B7D6B", "#FFB90F", "#8B008B","#7AC5CD"))+theme_bw() + 
  theme(axis.ticks = element_blank()) +theme(panel.grid=element_blank())+ theme(legend.title=element_blank()) #+geom_smooth(method = "loess", se=FALSE, color="black", aes(group=sponge_type))
ggpubr::ggarrange(p1,p2,ncol=2,nrow=1,labels=c("A","B"),common.legend = T)

#LM
OTU.full.i.WATER=OTU.full.i%>%filter(sponge_type=="WATER")
lm.WATER<-lm(OTU.full.i.WATER$incidence~OTU.full.i.WATER$hour)
summary(lm.WATER)# p-value =0.000662 ***,r2=0.4369

OTU.full.i.PINK=OTU.full.i%>%filter(sponge_type=="PINK")
lm.pink<-lm(OTU.full.i.PINK$incidence~OTU.full.i.PINK$hour)
summary(lm.pink)# p-value =0.4241,r2=-0.01691

OTU.full.i.CREAM=OTU.full.i%>%filter(sponge_type=="CREAM")
lm.CREAM<-lm(OTU.full.i.CREAM$incidence~OTU.full.i.CREAM$hour)
summary(lm.CREAM)# p-value =0.008214**,r2=0.2781
```

```{r incidence aov}
fit1<-aov(incidence~sponge_type,data=OTU.full.i)
summary(fit1)
#sponge_type  3  17.65   5.885   7.471 0.000182 ***
#remove black sponge
OTU.full.i2=OTU.full.i%>%filter(sponge_type!="BLACK")
OTU.full.i2$sponge_type=as.factor(as.character(OTU.full.i2$sponge_type))
leveneTest(incidence~sponge_type,data=OTU.full.i2)# 0.1842
fit2<-aov(incidence~sponge_type,data=OTU.full.i2)
summary(fit2)
#             Df  Sum Sq Mean Sq F value Pr(>F)
#sponge_type  2   0.67  0.3333   0.249   0.78 
plot(fit2, 2)

fit3<-aov(incidence~sponge_type+hours_fish,data=OTU.full.i2)
plot(fit3, 2)
summary(fit3)
#sponge_type  2   0.89   0.444   0.470 0.627803    
#hours_fish   6  25.94   4.323   4.567 0.000815 ***

par(mar=c(4,4,2,2))
interaction2wt(incidence~sponge_type*hours_fish,data=OTU.full.i2)

TukeyHSD(aov(incidence~sponge_type,data=OTU.full.i2)) 
#plot(TukeyHSD(aov(incidence~sponge_type,data=OTU.full.i)))
# library(ggsignif)
# library(gtools)
# compared_list=list()
# levels=levels(as.factor(OTU.full3$sponge_type))
# combinations=combinations(4,2,levels)
# for (i in 1:(length(combinations)/2)) {
#   compared_list[[i]]=combinations[i,]
# }
# ggplot(OTU.full3, aes(x=sponge_type, y=incidence)) + 
#   geom_boxplot() +
#   labs(title = "incidence t.test")+
#   geom_signif(comparisons = compared_list, test = t.test, step_increase = 0.2,map_signif_level=TRUE)+geom_jitter(width =0.3,size=1,height=0.0)
```
the beta diversity analysis base on no black data
```{r mvabund}
OTU.sam2=OTU.sam2 %>%filter(OTU.full.i$sponge_type!="BLACK" )
metadata.sam=OTU.full.i %>%filter(OTU.full.i$sponge_type!="BLACK")%>%select(hour,hours_fish,tank,sponge_type)
metadata.sam$sponge_type=as.factor(as.character(metadata.sam$sponge_type))
metadata.sam$tank=as.factor(as.character(metadata.sam$tank))

communityB.mvb <- mvabund(OTU.sam2)  
is.mvabund(communityB.mvb) # checking that the arthmvabund dataset is properly formatted
meanvar.plot(communityB.mvb,col=as.numeric(metadata.sam$sponge_type))
#plot(communityB.mvb ~ metadata.sam$sponge_type) # predictor must be a factor
nboot <- 999
#base model1 for sponge_type
modt1.mvb <- manyglm(communityB.mvb~sponge_type*hour*tank, family="negative.binomial",data=metadata.sam)
plot(modt1.mvb, res.type="pit.norm") 
anova(modt1.mvb,cor.type = "shrink", test = "score", show.time = "all", nBoot = nboot)
 
str(metadata.sam.c)
communityB.mvb <- mvabund(OTU.sam.c2) 
is.mvabund(communityB.mvb) 
modt3.mvb <- manyglm(communityB.mvb~hour*tank, family="binomial",data=metadata.sam.c)
plot(modt3.mvb)
anova(modt3.mvb,cor.type = "shrink", test = "score", show.time = "all", nBoot = nboot)

str(metadata.sam.p)
communityB.mvb <- mvabund(OTU.sam.p2) 
modt4.mvb <- manyglm(communityB.mvb~hour*tank, family="binomial",data=metadata.sam.p)
plot(modt4.mvb)
anova(modt4.mvb,cor.type = "shrink", test = "score", show.time = "all", nBoot = nboot)

str(metadata.sam.w)
communityB.mvb <- mvabund(OTU.sam.w2) 
modt5.mvb <- manyglm(communityB.mvb~hour*tank, family="binomial",data=metadata.sam.w)
plot(modt5.mvb)
anova(modt5.mvb,cor.type = "shrink", test = "score", show.time = "all", nBoot = nboot)

#            Res.Df Df.diff  score Pr(>score)   
#(Intercept)     20                             
#hour            19       1 12.304      0.004 **
#tank            17       2 21.912      0.002 **
#hour:tank       15       2  1.645      0.657   
```

plot pyramid: community change along time
```{r Pyramid}
brewer.pal(n =11, name = "RdYlBu")
col1=c("#E41A1C", "#FB8072", "#377EB8" )
col2=c("#E41A1C", "#FB8072", "#80B1D3" )
col3=c("#E41A1C", "#FB8072", "#ABD9E9" )

OTU.full2=OTU.full.i %>% filter(OTU.full.i$tank=="TANKC" & OTU.full.i$sponge_type =="WATER")
rowname=OTU.full2$name
OTU.full3=OTU.full2 %>% select(Amphiprion_ocellaris:Gramma_loreto)
exclude2 = which(colSums(OTU.full3) == 0)
OTU.full3=OTU.full3[,-exclude2]

OTU.full3=as.matrix(OTU.full3)
OTU.full3 <- reshape2::melt(OTU.full3)
names(OTU.full3) <- c("sample","species","reads")

ggplot(OTU.full3, aes(x =sample , y = reads, fill = species)) +
  geom_bar(stat = "identity", position='stack') +  
  theme_bw() + 
  theme(axis.ticks = element_blank()) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))+theme(panel.grid=element_blank()) +    
  theme(panel.border=element_blank())  + 
  theme(legend.title = element_blank(), legend.position = "left") + 
 scale_fill_manual(values=col3)+scale_y_continuous(limits=c(0,30),breaks=seq(0,30,5))
```

